---
title: "Reproducible Research: Peer Assessment 1 - By Shay (First Change)"
output: 
  html_document:
    keep_md: true
---


## Loading libraries and preprocessing the data
``` {r}
library(ggplot2)

data <- read.csv("./activity.csv")
```

### change date column to Date type 
``` {r}
data$date <- as.Date(as.character(data$date), "%Y-%m-%d")
```

## What is mean total number of steps taken per day?
``` {r}
hist(data$steps)
mean <- mean(data$steps,na.rm=TRUE)
median <- median(data$step,na.rm=TRUE)
```

#### Mean is `r mean`
#### Median is `r median`

## What is the average daily activity pattern?

#### Calculate the mean by interval ignoring NA's
``` {r echo=TRUE}
summary <- aggregate(data["steps"], by=data[c("interval")], FUN=mean, na.rm=TRUE)
```

#### Plot results
``` {r}
ggplot(summary, aes(x=interval,y=steps))+geom_line()
```

#### Select interval with highest average number of steps
``` {r}
summary <- summary[order(-summary$steps),]
max_steps_time <- summary[1,]$interval
```
#### Time witih highest number of steps is `r max_steps_time`. Indicates people starting their day at 8:35AM and going to work.

## Imputing missing values

####calculate number of rows w nas

``` {r}
NA_rows <- nrow(data) - nrow(data[complete.cases(data),]) 
```
####Number of rows with NAs is `r NA_rows`.

####Impute values for NAs using loess model based on interval time
``` {r}
Steps_Loess <- loess(steps ~ interval, span=.75, summary)
LoessPredicted <- predict(Steps_Loess,summary)
summary <- cbind(summary,LoessPredicted)
```

#### Here is representation of the model. It has obvious flaws but will use it for illustration purposes.
``` {r}
ggplot(summary, aes(x=interval,y=steps)) + 
geom_line(aes(interval,LoessPredicted)) +
geom_line(aes(interval,steps))
```

#### Create a seperate data set and impute NAs with loess model above.
``` {r}
data_imputed <- data
data_imputed$PredSteps <- predict(Steps_Loess,data_imputed$interval)
data_imputed$steps[is.na(data_imputed$steps)] <- data_imputed$PredSteps
```

#### Create histogram, mean, and median with imputed data
``` {r}
hist(data_imputed$steps)
imputed_mean <- mean(data_imputed$steps,na.rm=TRUE)
imputed_median <- median(data_imputed$step,na.rm=TRUE)
```

#### Mean is `r imputed_mean` compared to the original value of `r mean`
#### Median is `r imputed_median`. compared to the original value of `r median`

## Are there differences in activity patterns between weekdays and weekends?

``` {r}
data_imputed$weekdays <- as.factor(weekdays(data_imputed$date))
data_imputed$day_type <- data_imputed$weekdays
levels(data_imputed$day_type) <- c("weekday","weekday","weekend","weekend","weekday","weekday","weekday")

summary_imputed <- aggregate(data_imputed["steps"], by=data_imputed[c("day_type","interval")], FUN=mean)
```


#### Plot the summary as before
``` {r}
ggplot(summary_imputed, aes(x=interval,y=steps, group=day_type)) +
facet_grid(day_type ~ .) +
geom_line() 
```


